Appendix

### 2025-10-28 manuscript_rmarkdown/manuscript.Rmd
Figure 5 without outlier-rate.

The null-hypothesis should be rejected at significance level $\alpha$ if 
$\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \zeta, \text{critical}}$,
with $\zeta$ the population estimate of the rate of outliers from a normal
distribution, to be defined in the [Simulation](#simulation) section.

@@

The critical test statistic for the empirical central normality test depends on 
the significance level $\alpha$, sequence length $n$, central portion $\kappa$
and estimated outlier rate $\zeta$.
Critical test statistic values were set through simulation, as described below.

For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
we randomly sampled $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
Outliers were inserted by randomly replacing elements of an ordered sequence with probability 
$\zeta \in \left\{0.0, 0.01, 0.02, 0.05, 0.10, 0.15, 0.20 \right\}$. 
Outlier values were set according to @Tukey1977-xm, as follows.
Let $x^{*} \sim U\left(-2, 2\right)$.
Then the corresponding outlier value was:

For each sequence we then computed $\tau_{\kappa}$ for $\kappa \in \left\{0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.

@@

Figure \ref{fig:empirical-central-normality-kappa} shows the critical test static
$\tau_{\alpha = 0.05, n, \kappa, \zeta}$
as a function of $n$ for different values of $\kappa$ and $\zeta$.
With decreasing $\kappa$, the test statistic curve decreases.
The curves for $\kappa = 0.60$, $\kappa = 0.70$ and $\kappa = 0.80$ are similar,
whereas curves for $\kappa \geq 0.90$ were affected by outliers.

We use $\kappa = 0.80$ and $\zeta = 0.10$ for the empirical central normality tests 
in the remainder of the manuscript, unless noted otherwise.
Critical statistic values of the empirical central normality test for 
$\kappa = 0.80$ and $\zeta = 0.10$ are shown in Table 
\ref{tab:empirical-central-normality-critical-statistic}.

```{r empirical-central-normality-kappa, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa, \\zeta}$ of the empirical ",
  "central normality test as function of $n$ for several values of central portion $\\kappa$ ",
  "and estimated outlier rates $\\zeta$. ",
  "emp: empirical"
)

.plot_test_statistic_centrality(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir
)

@@

To assess type I error rates for the empirical central normality test 
($\kappa = 0.80$ and $\zeta \in \left\{0.00, 0.05, 0.10\right\}$) and compare
these to the Shapiro-Wilk test for normality, we randomly drew another 10000 sequences from a
standard normal distribution ($\mathcal{N}(0,1)$ for each 
$n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.
Each sequence was additionally randomly populated with outliers at 5% and 10% rates, as earlier.

@@

In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical
central normality test to assess central normality of features that are composed
of a mixture of samples drawn from two normal distributions ($n = 100$ each).
With increased separation of the underlying normal distributions, the probability of the
feature being centrally normal decreases, as expected.

```{r empirical-central-normality-examples, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Bi-modal distributions and empirical central normality test results. ",
  "The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) ",
  "drawn from normal distributions that are offset by a distance $d$. ",
  "We use the empirical centrally normality test to compute the probability for the hypothesis that ",
  "the distribution is centrally normal. As may be observed, with increasing offset $d$ ",
  "the probability that the feature is centrally normal decreases. ",
  "Quantile-quantile plots are drawn below each distribution."
)

@@

@@

To determine the weighting function parameters for each of the nine combinations,
$m_d=500$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of sequences with between $50$ and $1000$ elements each.
Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.

>>>>

The null-hypothesis should be rejected at significance level $\alpha$ if 
$\tau_{n, \kappa} \geq \tau_{\alpha, n, \kappa, \text{critical}}$.
Critical test statistic values are determined in 
the [Simulation](#simulation-ecn-test) section.

@@

## Empirical central normality test {#simulation-ecn-test}

The critical test statistic for the empirical central normality test depends on 
the significance level $\alpha$, sequence length $n$ and central portion $\kappa$.
Critical test statistic values were set through simulation, as described below.

For each number of samples $n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 4.0000 \right \} \right\}$
we randomly sampled $m_d = 30000$ ordered sequences $\mathbf{X}$ from $\mathcal{N}(0,1)$.
For each sequence we then computed $\tau_{\kappa}$ for 
$\kappa \in \left\{0.50, 0.60, 0.70, 0.80, 0.90, 0.95, 1.00\right\}$.

@@

Figure \ref{fig:empirical-central-normality-combined} shows the critical test statistic
$\tau_{\alpha, n, \kappa}$ as a function of $n$ for different values of $\kappa$ and $\alpha$. 
For larger central portion $\kappa$, critical test statistic values increase.
Similarly, critical test statistic values increase for lower significance levels.

```{r empirical-central-normality-combined, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha = 0.05, n, \\kappa}$ of the empirical ",
  "central normality test as function of $n$ for several values of central portion $\\kappa$ (left) ",
  "and several values of significance level $\\alpha$ (right)."
)

.plot_test_statistic_combined(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir
)

@@

To assess type I error rates for the empirical central normality test and compare
these to the Shapiro-Wilk test for normality, we randomly drew another 10000 sequences from a
central normal distribution for each 
$n \in \left\{\lfloor 10^\nu \rfloor | \nu \in \left\{0.7500, 0.8125, \ldots, 3.0000 \right \} \right\}$.

Each sequence was created as follows: first $n$ values are sampled from $\mathcal{N}(0, 1)$.
Then, the elements outside the central portion of the sequence (i.e. $\mathbf{X} \setminus \mathbf{X}_{\text{c}}$) are replaced.
Elements with $pi < \frac{1-\kappa}{2}$ are replaced with values sampled from $\mathcal{U}(-10,\min(\mathbf{X}_{\text{c}}))$,
and elements with $p_i > \frac{1 + \kappa}{2}$ are replaced with values sampled from $\mathcal{U}(\max(\mathbf{X}_{\text{c}}), 10)$.

@@

In Figure \ref{fig:empirical-central-normality-examples} we apply the empirical
central normality test (with $\kappa = 0.80$) to assess central normality of features that are composed
of a mixture of samples drawn from two normal distributions ($n = 100$ each).
With increased separation of the underlying normal distributions, 
the test's p-value decreases, as expected.

```{r empirical-central-normality-examples, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Bi-modal distributions and empirical central normality test results. ",
  "The feature (black) is a mixture of two identical sample sets (blue and orange, $n = 100$) ",
  "drawn from normal distributions that are offset by a distance $d$. ",
  "The empirical central normality ($\\kappa = 0.80$) is used to test the hypothesis that the mixture is sampled from ",
  "a central normal distribution. As may be observed, with increasing offset $d$ ",
  "the test's p-value decreases. ",
  "Quantile-quantile plots are drawn below each distribution."
)

@@

We use $\kappa = 0.80$ for the empirical central normality tests 
in the remainder of the manuscript, unless noted otherwise.
Critical statistic values of the empirical central normality test for 
$\kappa = 0.80$ are shown in Table 
\ref{tab:empirical-central-normality-critical-statistic}.

@@

To determine the weighting function parameters for each of the nine combinations,
$m_d=500$ sequences $\mathbf{X}_i$ ($i \in \{1, 2, \ldots, m_d\}$) were randomly 
drawn from randomly parametrised asymmetric generalised normal distributions.
Each distribution was parametrised with a random skewness parameter $\alpha \sim U\left(0.01, 0.99\right)$ 
and shape parameter $\beta \sim U\left(1.00, 5.00 \right)$.
Location and scale parameters were set as $\mu = 0$ and $\sigma = 1$, respectively.
To form each sequence $\mathbf{X}_i$, $n = \lceil 10^\gamma \rceil$ instances were then randomly drawn, 
with $\gamma \sim U\left(\log_{10}50, 3\right)$, resulting in a set of sequences with between $50$ and $1000$ elements each.
Outlier values were then drawn to randomly replace 10 percent of the elements of $\mathbf{X}_i$.
These values were set according to @Tukey1977-xm, as follows.
Let $x^{*} \sim U\left(-2, 2\right)$.
Then the corresponding outlier value was:

### 2025-10-28 manuscript_rmarkdown/manuscript_appendix.Rmd
Figure 5 without outlier-rate.

Critical values for the empirical central normality test are
found in Table \ref{tab:central-normality-critical-statistic} and Table 
\ref{tab:empirical-central-normality-critical-statistic}, respectively.

```{r include = FALSE, eval = FALSE}
# This generates the table shown below.
.get_test_statistic_lookup_table_manuscript(
  manuscript_dir = manuscript_dir,
  kappa_lookup = 0.80,
  outlier_rate_lookup = 0.10
)

@@

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80, \zeta = 0.10}$ for
central normality at $\kappa = 0.80$ and outlier rate $\zeta = 0.10$, as a function of significance 
level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

@@

\midrule
    5 & 210.38 & 82.23 & 59.47 & 48.68 & 38.71 & 29.28 & 19.75 & 13.71 & 11.15 \\
   10 &  43.22 & 33.68 & 29.26 & 26.04 & 22.96 & 19.62 & 14.37 & 10.50 &  8.89 \\
   20 &  31.31 & 25.35 & 22.61 & 20.58 & 18.23 & 15.64 & 11.51 &  8.61 &  7.49 \\
   50 &  19.97 & 16.53 & 14.83 & 13.33 & 11.77 & 10.11 &  7.60 &  5.85 &  5.16 \\
  100 &  14.01 & 11.57 & 10.30 &  9.31 &  8.31 &  7.22 &  5.49 &  4.29 &  3.78 \\
  200 &   9.87 &  8.15 &  7.39 &  6.73 &  6.02 &  5.28 &  4.06 &  3.16 &  2.80 \\
  500 &   6.30 &  5.34 &  4.87 &  4.49 &  4.08 &  3.60 &  2.80 &  2.17 &  1.92 \\
 1000 &   4.72 &  4.02 &  3.70 &  3.43 &  3.15 &  2.81 &  2.21 &  1.72 &  1.51 \\
 2000 &   3.61 &  3.18 &  2.96 &  2.78 &  2.57 &  2.32 &  1.88 &  1.49 &  1.31 \\
 5000 &   2.80 &  2.51 &  2.37 &  2.25 &  2.12 &  1.96 &  1.66 &  1.38 &  1.24 \\
10000 &   2.41 &  2.20 &  2.10 &  2.02 &  1.92 &  1.81 &  1.60 &  1.39 &  1.28 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}

```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.80, \\zeta}$ of the empirical ",
  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
  "emp: empirical"
)

.plot_test_statistic_tau(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir
)

>>>>

Critical values for the empirical central normality test at $\kappa = 0.80$ are
found in Table \ref{tab:empirical-central-normality-critical-statistic}.

```{r include = FALSE, eval = FALSE}
# This generates the table shown below.
.get_test_statistic_lookup_table_manuscript(
  manuscript_dir = manuscript_dir,
  kappa_lookup = 0.80
)

@@

\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.80}$ for
central normality at $\kappa = 0.80$, as a function of significance 
level $\alpha$ and number of instances $n$. The values in this table should be divided by $100$ before use.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

@@

\midrule
    5 & 200.88 & 90.25 & 64.46 & 49.04 & 36.07 & 27.70 & 19.28 & 13.41 & 10.90 \\
   10 &  45.90 & 31.78 & 27.12 & 24.14 & 21.23 & 18.30 & 13.68 & 10.20 &  8.71 \\
   20 &  28.88 & 22.53 & 19.96 & 18.01 & 15.99 & 13.93 & 10.67 &  8.26 &  7.27 \\
   50 &  16.86 & 14.10 & 12.66 & 11.55 & 10.28 &  8.95 &  7.00 &  5.53 &  4.91 \\
  100 &  12.41 & 10.00 &  9.02 &  8.21 &  7.36 &  6.45 &  5.02 &  4.00 &  3.58 \\
  200 &   8.65 &  7.05 &  6.37 &  5.80 &  5.22 &  4.57 &  3.57 &  2.84 &  2.54 \\
  500 &   5.54 &  4.47 &  4.03 &  3.69 &  3.30 &  2.90 &  2.27 &  1.81 &  1.63 \\
 1000 &   3.88 &  3.17 &  2.86 &  2.60 &  2.34 &  2.05 &  1.61 &  1.28 &  1.15 \\
 2000 &   2.79 &  2.24 &  2.03 &  1.84 &  1.66 &  1.45 &  1.14 &  0.91 &  0.82 \\ 
 5000 &   1.75 &  1.41 &  1.27 &  1.17 &  1.05 &  0.92 &  0.72 &  0.58 &  0.52 \\
10000 &   1.21 &  1.00 &  0.90 &  0.82 &  0.74 &  0.65 &  0.51 &  0.41 &  0.36 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
